### 2023-04-15 blog.qmd
Update blog.qmd

---
title: "One Day"
listing:
  contents: posts
  sort: "date desc"
  type: default
  categories: true
  sort-ui: true
  filter-ui: true
  fields: [date, title, reading-time, description, categories]
page-layout: full
title-block-banner: false
---

# Series

##### [rtables-series](rtables-series.html)
This series contains a great deal of tips, tricks and packages
that you can use to level up your `rtables` game.

##### [manipulate-series](manipulate-series.html)
This series contains a great deal of tips, tricks of `dplyr`