### 2023-04-15 posts/rtables/customize_cell.qmd
Create customize_cell.qmd

---
title: 'Customize the format of rcell'
date: '2023-04-06'
author: Qing Zou
categories: ['R', 'rtables']
description: 'Construct a cell value and associate formatting'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(rtables)
library(formatters)
```

## rcell

__Description__

Construct a cell value and associate formatting, labeling, indenting, and column spanning information with it.

__Usage__

```{r rcell1}
#| echo: true
#| eval: false
rcell(
x,
format = NULL,
colspan = 1L,
rcell 103
label = NULL,
indent_mod = NULL,
footnotes = NULL,
align = NULL,
format_na_str = NULL
)
```

__Arguments__

  |
 -|:--
 __x__| ANY. Cell value
 __format__| if FUN does not return a formatted rcell then the format is applied
 __...__|  further arguments passed to rcell

### Real case

If we have a calculation `1/4*100`, the result `25` and the decimal is `2`.
the avaiable format in rtables is `xx.xx`

```{r rcell2}
#| echo: true
#| eval: true
rcell(
  x = 25,
  format = 'xx.xx'
)
```

What if we want a format like `1 (25.00)`? how can we customize the format in rcell as expected?
Let's go through the function of `sprintf` firstly. the `sprintf` can returns a character vector containing a formatted combination of text and variable values.

1. print the `25` to `25.00`

```{r rcell3}
#| echo: true
#| eval: true
sprintf("%.2f", 25)
```
For the given format, `%`: it's like a place holder for the passed value, `f`: Double precision value, see more on [sprintf](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sprintf)

2. update the format and passed value

```{r rcell4}
#| echo: true
#| eval: true
sprintf("%.0f %.2f", 1, 25)
```

```{r rcell5}
#| echo: true
#| eval: true
sprintf("%.0f (%.2f)", 1, 25)
```

---

In order to apply this format on other values, can we separate the formatting `%.0f (%.2f)` and the passed value `1` and `25`?

there are two functions from package `formatters`

`sprintf_format`: Specify text format

`format_value`: Converts a value into a string using the format information

```{r rcell6}
#| echo: true
#| eval: true
fmtfun <- formatters::sprintf_format("%.0f (%.2f)")

format_value(x = c(1, 25),
             format = fmtfun)
```

As for the `rcell` and  its argument `x` and `format`, we can think it is a wrapper around the `format_value`

```{r rcell7}
#| echo: true
#| eval: true
rcell(
  x = c(1, 25),
  format = formatters::sprintf_format("%.0f (%.2f)")
)
```

Then, we can also easily customize the result of `1/4*100` as `1/4 (25.00)`

```{r rcell8}
#| echo: true
#| eval: true
rcell(
  x = c(1, 4, 25),
  format = formatters::sprintf_format("%.0f/%.0f (%.2f)")
)
```

### Create format helper function

we can further more wrap this pair of funcions into two helper function for better usage.

```{r rcell9}
#| echo: true
#| eval: true

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param decimal Number of decimal to display
#' @param type to decide if the denominator n/N(%) in the result
#' @keywords sprintf_format
#'
#' @return
sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- "/%.0f"
  }

  if (n == 0) {
    fmt <- sprintf_format(paste0("%.0f", showN))
  } else if (n == N) {
    # if pct == 100%, then show (100) instead of (100.0)
    fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
  } else {
    fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
  }

  return(fmt)
}

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param type to decide if the denominator n/N(%) in the result
#'
#' @keywords format_values
#' @example
#' rcell(x = format_values_helper(1, 10),
#'       format = sprintf_format_helper(1, 10))
#'
#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
#'
#' rcell(x = format_values_helper(10, 10),
#'       format = sprintf_format_helper(10, 10))
#'
#' @return
format_values_helper <- function(n, N, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- N
  }
  if (n == 0) {
    value <- c(0, showN)
  } else {
    value <- c(n, showN, n / N * 100)
  }

  return(value)
}
```

then, we can apply various formatter during printing values in `rcell`.

```{r rcell10}
#| echo: true
#| eval: true
rcell(x = format_values_helper(1, 10),
      format = sprintf_format_helper(1, 10))
```

```{r rcell11}
#| echo: true
#| eval: true

rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
      format = sprintf_format_helper(1, 10, type = "n/N(%)"))

```

```{r rcell12}
#| echo: true
#| eval: true

rcell(x = format_values_helper(10, 10),
      format = sprintf_format_helper(10, 10))
```

```{r rcell13}
#| echo: true
#| eval: true

rcell(x = format_values_helper(0, 10, type = "n/N(%)"),
      format = sprintf_format_helper(0, 10, type = "n/N(%)"))
```