### 2023-04-15 posts/rtables/multiple_levels_calculation_1.qmd
Create multiple_levels_calculation_1.qmd

---
title: 'Mutiple level calculation 1'
date: '2023-04-07'
author: Qing Zou
categories: ['R', 'rtables']
description: 'Explore how to handle the mutiple calculation by rtables'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell3sd45}
#| echo: false
#| eval: true

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param decimal Number of decimal to display
#' @param type to decide if the denominator n/N(%) in the result
#' @keywords sprintf_format
#'
#' @return
sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- "/%.0f"
  }

  if (n == 0) {
    fmt <- sprintf_format(paste0("%.0f", showN))
  } else if (n == N) {
    # if pct == 100%, then show (100) instead of (100.0)
    fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
  } else {
    fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
  }

  return(fmt)
}

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param type to decide if the denominator n/N(%) in the result
#'
#' @keywords format_values
#' @example
#' rcell(x = format_values_helper(1, 10),
#'       format = sprintf_format_helper(1, 10))
#'
#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
#'
#' rcell(x = format_values_helper(10, 10),
#'       format = sprintf_format_helper(10, 10))
#'
#' @return
format_values_helper <- function(n, N, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- N
  }
  if (n == 0) {
    value <- c(0, showN)
  } else {
    value <- c(n, showN, n / N * 100)
  }

  return(value)
}

#' Display rtables in html
#'
#' @param table a rtables object
#' @param to_html export to HTML. Default TRUE
#'
#' @return a html table
#' @export
table_to_html <- function(table, to_html = TRUE) {
  stopifnot(!is.null(table), class(table) %in% c("TableTree", "ElementaryTable"))

  extra_css <- "
  .tabwid {overflow-x: scroll; display: block;}
  .tabwid table td p {min-width: 170px !important;}
  .tabwid table td:first-child > p {min-width: 200px !important;}
  .tabwid table tr:hover {background-color: #ffff99 !important;}
  "
  flextable::set_flextable_defaults(
    font.family = ifelse(to_html, "Source Sans Pro", "Verdana"),
    extra_css = extra_css
  )

  ft <- rtables::tt_to_flextable(table) %>%
    flextable::border_remove() %>%
    flextable::bold(bold = TRUE, part = "header") %>%
    flextable::hline_top(part = "body") %>%
    flextable::hline_bottom(part = "body") %>%
    flextable::autofit()

  if (to_html) {
    flextable::htmltools_value(ft)
  } else {
    ft
  }
}
```

```{r cal1}
library(rtables)
library(formatters)
library(dplyr)
```

## Multiple lelvels calculation

`Multipile levels` calculation for categorical variable is the most common cases in Table, e.g. `PCSA`, `Exposure tables`.

In this section, we will explore how to produce these output in `rtables`.

Below is the test data:

```{r cal2}
#| echo: true
#| eval: true
df <- data.frame(SUBJID = c(rep('S1', 5), rep("S2", 4), rep("S3", 2)),
                 ASEQ = c(1:5, 1:4, 1:2),
                 GROUP = factor(c(rep('A', 5), rep("B", 4), rep('A', 2)), levels = c("A", "B")),
                 PARAMCD = c('TEST_1', 'TEST_1', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_2', 'TEST_1', 'TEST_2'),
                 CRIT = c('LEVEL_2_1','LEVEL_2_1', 'LEVEL_2_2', NA_character_, 'LEVEL_2_a', 'LEVEL_2_1', NA_character_, 'LEVEL_2_a', 'LEVEL_2_b', NA_character_, 'LEVEL_2_b'))
```
```{r cal2sd }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

Now, we need to take the `PARAMCD` as 1st level, and the `CRIT` as 2nd level to calculate the frequency of each `TEST`, it will be a table like PCSA.

# Method 1

the `rtables::analyze()` will firstly be used in this case.

To use this function, the data should be transformed from `long strcture(BDS)` to `wide`.

Steps:

1. take action on setting up for the `missing value` and the `factor`:

 - fill the missing value as `Missing_value_flag`, which will be subsequently handled in `rtables::analyze()`
 - use function `tidyr::pivot_wider` implement the transformation, the new columns are used as the analyzed variables `vars` of `rtables::analyze()`
 - `factor` the values by given level values, the levels is the format of output

```{r cal2df}
#| echo: true
#| eval: true

#
df <- df %>% mutate(across(CRIT, ~ifelse(.=="" | is.na(.), "Missing_value_flag", as.character(.))))

df1 <- df %>% tidyr::pivot_wider(id_cols = c("SUBJID", 'ASEQ', 'GROUP'),
                                 names_from = "PARAMCD",values_from = "CRIT")

#factor the data by mock
fmt <- list(
  'TEST_1' = c('LEVEL_2_1' = '1 of level 2',
               'LEVEL_2_2' = '2 of level 2',
               'LEVEL_2_3' = '3 of level 2'),
  "TEST_2" = c('LEVEL_2_a' = 'a of level 2',
              'LEVEL_2_b' = 'b of level 2',
              'LEVEL_2_c' = 'c of level 2')
)

fmt <- purrr::map(fmt, ~append(., c("Missing_value_flag" = "Missing_value_flag")))

df1 <- df1 %>%
  mutate(across(paste0('TEST_', 1:2), ~factor(fmt[[cur_column()]][.x],
                                              levels = c(fmt[[cur_column()]])), .names = "{.col}_fmt"))
```
```{r cal2g }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

2. The dataframe `df1` is ready for output. now, the `layout` and `key values` of table will be built in the `rtables::analyze()`.

 - the header and overall column are generated by  `split_cols_by` and `add_overall_col`
 - in the `rtables::analyze()`:
   - `vars` and `var_labels` are the variable names and labels
   - in the `afun`, we create an analysis function, in which the arguments `df`, `.N_col`, `.var`, `.N_row` are created by rtables itself. the other aruments are customized will be passed by `extral_args`
     - the `df` is current available(`non NA`) data with splited group, that's why we firstly fill the missing data `Missing_value_flag`. By filling
       a non-NA value, we can handle the missing value as users expected.
     - you will see the customized arguments `missing_rm` and `missing_label`, if user hope hide or remove the missing values.
     - the `n_unique_keys` and `N_unique_keys` indicate that when counting the number of n/N, which records should be used.
     - `add_Number` means whether add an row `Number: N`
     - All the values of these customized arguments are passed by the `extra_args`, it's the greatest way to add the flexiabiliy for the tabulation.

```{r cal2as}
#| echo: true
#| eval: true

lyt <- basic_table() %>%
  split_cols_by("GROUP") %>%
  add_overall_col("All") %>%
  rtables::analyze(vars = c('TEST_1_fmt', 'TEST_2_fmt'),
                   var_labels = c('This is test 1', 'This is test 2'),
                   afun = function(df, .N_col, .var, .N_row,
                                  format = 'n(%)',
                                  decimal = 1,
                                  n_unique_keys = NULL,
                                  N_unique_keys = NULL,
                                  show_N_item = FALSE,
                                  add_Number = FALSE,
                                  missing_rm = TRUE,
                                  missing_label =NULL,
                                  ...) {

                    if(missing_rm){
                      df <- df %>% filter(!(!!as.name(.var) %in% "Missing_value_flag"))
                    }

                    df_n <- df
                    if(!is.null(n_unique_keys)){
                      df_n <- df %>% distinct(across(c(n_unique_keys, .var)), .keep_all = TRUE)
                    }

                    x <- df_n[[.var]]

                    if(!is.null(N_unique_keys)){
                      cus_N_col <- df %>% distinct(across(N_unique_keys)) %>% nrow()
                    }else{
                      cus_N_col <- .N_col
                    }

                    lst_body <- lapply(
                      as.list(table(x)),
                      function(xi) {
                        rcell(format_values_helper(xi, cus_N_col, type = format),
                              format = sprintf_format_helper(xi, cus_N_col, 1, type = format),
                              indent_mod = 1L
                        )
                      }
                    )

                    if(is.null(missing_label)){
                      if("Missing_value_flag" %in% names(lst_body)){
                        lst_body <- purrr::list_modify(lst_body, "Missing_value_flag" = NULL)
                      }
                    }else{
                      lst_body <- lst_body %>%
                        setNames(gsub("Missing_value_flag", missing_label, names(.)))

                    }

                    if(add_Number){
                      lst_body <- append(setNames(list(rcell(cus_N_col,
                                                            format = "xx",
                                                            indent_mod = 1L)),
                                                  'Number'), lst_body)
                    }
                    in_rows(
                      .list = c(
                        lst_body,
                        # add empty row, in .list obj
                        " " = rcell("", format = "xx")
                      )
                    )
                  },
          extra_args = list(format = 'n/N(%)',
                            n_unique_keys = c("SUBJID"),
                            N_unique_keys = c("SUBJID"),
                            add_Number = FALSE,
                            missing_rm = FALSE,
                            missing_label = NULL)
          # ,
          # show_labels = "hidden"
  )
```
3. finally, fit the `layout` and `df1` into the `rtables::build_table`

```{r cal2ad}
#| echo: true
#| eval: true
col_counts_df <- df1 %>%
  distinct(SUBJID, GROUP)
col_counts <- table(col_counts_df$GROUP)
col_counts <- c(col_counts, sum(col_counts))

tbl <- lyt %>% rtables::build_table(df1, col_counts = col_counts)
```

```{r cap1}
#| echo: false
#| eval: true
my_caption <- ''
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(tbl, to_html = FALSE)
```