### 2023-04-15 posts/rtables/multiple_levels_calculation_2.qmd
Create multiple_levels_calculation_2.qmd

---
title: 'Mutiple level calculation 2'
date: '2023-04-09'
author: Qing Zou
categories: ['R', 'rtables']
description: 'Wrap the analysis function'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell3sd45}
#| echo: false
#| eval: true

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param decimal Number of decimal to display
#' @param type to decide if the denominator n/N(%) in the result
#' @keywords sprintf_format
#'
#' @return
sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- "/%.0f"
  }

  if (n == 0) {
    fmt <- sprintf_format(paste0("%.0f", showN))
  } else if (n == N) {
    # if pct == 100%, then show (100) instead of (100.0)
    fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
  } else {
    fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
  }

  return(fmt)
}

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param type to decide if the denominator n/N(%) in the result
#'
#' @keywords format_values
#' @example
#' rcell(x = format_values_helper(1, 10),
#'       format = sprintf_format_helper(1, 10))
#'
#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
#'
#' rcell(x = format_values_helper(10, 10),
#'       format = sprintf_format_helper(10, 10))
#'
#' @return
format_values_helper <- function(n, N, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- N
  }
  if (n == 0) {
    value <- c(0, showN)
  } else {
    value <- c(n, showN, n / N * 100)
  }

  return(value)
}

#' Display rtables in html
#'
#' @param table a rtables object
#' @param to_html export to HTML. Default TRUE
#'
#' @return a html table
#' @export
table_to_html <- function(table, to_html = TRUE) {
  stopifnot(!is.null(table), class(table) %in% c("TableTree", "ElementaryTable"))

  extra_css <- "
  .tabwid {overflow-x: scroll; display: block;}
  .tabwid table td p {min-width: 170px !important;}
  .tabwid table td:first-child > p {min-width: 200px !important;}
  .tabwid table tr:hover {background-color: #ffff99 !important;}
  "
  flextable::set_flextable_defaults(
    font.family = ifelse(to_html, "Source Sans Pro", "Verdana"),
    extra_css = extra_css
  )

  ft <- rtables::tt_to_flextable(table) %>%
    flextable::border_remove() %>%
    flextable::bold(bold = TRUE, part = "header") %>%
    flextable::hline_top(part = "body") %>%
    flextable::hline_bottom(part = "body") %>%
    flextable::autofit()

  if (to_html) {
    flextable::htmltools_value(ft)
  } else {
    ft
  }
}
```

```{r cal1}
library(rtables)
library(formatters)
library(dplyr)
```

```{r cal2}
#| echo: false
#| eval: true
df <- data.frame(SUBJID = c(rep('S1', 5), rep("S2", 4), rep("S3", 2)),
                 ASEQ = c(1:5, 1:4, 1:2),
                 GROUP = factor(c(rep('A', 5), rep("B", 4), rep('A', 2)), levels = c("A", "B")),
                 PARAMCD = c('TEST_1', 'TEST_1', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_2', 'TEST_1', 'TEST_2'),
                 CRIT = c('LEVEL_2_1','LEVEL_2_1', 'LEVEL_2_2', NA_character_, 'LEVEL_2_a', 'LEVEL_2_1', NA_character_, 'LEVEL_2_a', 'LEVEL_2_b', NA_character_, 'LEVEL_2_b'))
```

```{r cal2df}
#| echo: false
#| eval: true

#
df <- df %>% mutate(across(CRIT, ~ifelse(.=="" | is.na(.), "Missing_value_flag", as.character(.))))

df1 <- df %>% tidyr::pivot_wider(id_cols = c("SUBJID", 'ASEQ', 'GROUP'), names_from = "PARAMCD",values_from = "CRIT")

#factor the data by mock
fmt <- list(
  'TEST_1' = c('LEVEL_2_1' = '1 of level 2',
               'LEVEL_2_2' = '2 of level 2',
               'LEVEL_2_3' = '3 of level 2'),
  "TEST_2" = c('LEVEL_2_a' = 'a of level 2',
              'LEVEL_2_b' = 'b of level 2',
              'LEVEL_2_c' = 'c of level 2')
)

fmt <- purrr::map(fmt, ~append(., c("Missing_value_flag" = "Missing_value_flag")))

df1 <- df1 %>%
  mutate(across(paste0('TEST_', 1:2), ~factor(fmt[[cur_column()]][.x], levels = c(fmt[[cur_column()]])), .names = "{.col}_fmt"))

col_counts_df <- df1 %>%
  distinct(SUBJID, GROUP)
col_counts <- table(col_counts_df$GROUP)
col_counts <- c(col_counts, sum(col_counts))
```

## Wrap the analysis function

The function is too long and it's not convient mixed with the layout of rtbales.
For better user experience, we can create a function `summarise_vars_categorical`

1. wrap the analysis function as `summarise_vars_categorical_afun`
2. wrap the `summarise_vars_categorical_afun` and the `extral_args` within the `summarise_vars_categorical`

Then user can passed all the analysis variables and customized argument by a single function `summarise_vars_categorical`.

```{r a23s}
#| echo: true
#| eval: true
#| message: FALSE

# wrap the analysis function
summarise_vars_categorical_afun <- function(df, .N_col, .var, .N_row,
                                            format = 'n(%)',
                                            decimal = 1,
                                            n_unique_keys = NULL,
                                            N_unique_keys = NULL,
                                            show_N_item = FALSE,
                                            add_Number = FALSE,
                                            missing_rm = TRUE,
                                            missing_label = NULL,
                                            indent_mod = 0L,
                                            ...) {

  if(missing_rm){
    df <- df %>% filter(!(!!as.name(.var) %in% "Missing_value_flag"))
  }

  df_n <- df
  if(!is.null(n_unique_keys)){
    df_n <- df %>% distinct(across(c(n_unique_keys, .var)), .keep_all = TRUE)
  }
  x <- df_n[[.var]]
  if(!is.null(N_unique_keys)){
    cus_N_col <- df %>% distinct(across(N_unique_keys)) %>% nrow()
  }else{
    cus_N_col <- .N_col
  }

  lst_body <- lapply(
    as.list(table(x)),
    function(xi) {
      rcell(format_values_helper(xi, cus_N_col, type = format),
            format = sprintf_format_helper(xi, cus_N_col, 1, type = format),
            indent_mod = indent_mod
      )
    }
  )
  if(add_Number){
    lst_body <- append(setNames(list(rcell(cus_N_col,
                                           format = "xx",
                                           indent_mod = indent_mod)), 'Number'), lst_body)
  }

  if(is.null(missing_label)){
    if("Missing_value_flag" %in% names(lst_body)){
      lst_body <- purrr::list_modify(lst_body, "Missing_value_flag" = NULL)
    }
  }else{
    lst_body <- lst_body %>%
      setNames(gsub("Missing_value_flag", missing_label, names(.)))

  }

  in_rows(
    .list = c(
      lst_body,
      # add empty row, in .list obj
      " " = rcell("", format = "xx")
    )
  )
}

# wrap the summarise_vars_categorical_afun and the extral_args
summarise_vars_categorical <- function(lyt,
                                       vars,
                                       var_labels,
                                       format = 'n(%)',
                                       decimal = 1,
                                       n_unique_keys = NULL,
                                       N_unique_keys = NULL,
                                       show_N_item = FALSE,
                                       add_Number = FALSE,
                                       missing_rm = TRUE,
                                       missing_label = NULL,
                                       show_labels='visible',
                                       indent_mod = 0L, ...){
  rtables::analyze(lyt, vars, var_labels,
                   afun = summarise_vars_categorical_afun,
                   extra_args = list(format = format,
                                     decimal = decimal,
                                     n_unique_keys = n_unique_keys,
                                     N_unique_keys = N_unique_keys,
                                     show_N_item = show_N_item,
                                     add_Number = add_Number,
                                     missing_rm = missing_rm,
                                     missing_label = missing_label),
                   show_labels = show_labels, indent_mod = indent_mod)
}

lyt <- basic_table() %>%
  split_cols_by("GROUP") %>%
  add_overall_col("All") %>%
  summarise_vars_categorical(vars = c('TEST_1_fmt', 'TEST_2_fmt'),
                             var_labels = c('This is test 1', 'This is test 2'),
                             format = 'n/N(%)',
                             n_unique_keys = c("SUBJID"),
                             N_unique_keys = c("SUBJID"),
                             add_Number = TRUE,
                             missing_rm = FALSE

  )

tbl <- lyt %>% rtables::build_table(df1, col_counts = col_counts)
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(tbl, to_html = FALSE)
```