### 2023-04-15 posts/manipulate/dplyr_cross.qmd
Create dplyr_cross.qmd

---
title: 'Derive multiple columns based on across'
date: '2023-04-08'
author: Qing Zou
categories: ['R', 'dplyr', 'across']
description: 'When we want to add new columns and have to iterate...'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(dplyr)
```

## Reale Case

Suppose we this data with columns `num_var1_1`, `num_var1_99` and `denom1_1`, `denom1_99`

```{r rcell15}
#| echo: true
#| eval: true
test_data <- data.frame(num_var1_1 = c(1, 4),
                        num_var1_99 = c(2, 6),
                        denom1_1 = c(2, 5),
                        denom1_99 = c(3, 1))
```

what we want to do is to add two new columns: the new column equal to the sum of `num_var_x` and `denom1_y`.
Just fit the data into the function `sum`

```{r rcell116}
#| echo: true
#| eval: true

sum_c <- function(x, y){
  ifelse(y > 2, x+y+10, x+y)
}

test_data1 <- test_data %>%
             mutate(var1_1 = sum_c(num_var1_1, denom1_1),
                    var1_99 = sum_c(num_var1_99, denom1_99))
```

```{r rcell11612 }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(test_data1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

the code seems not readable.

## dplyr::across()

we can use the across() to easily apply the functions to the selected columns.

__Description__

`across()` makes it easy to apply the same transformation to multiple columns, allowing you to use `select()` semantics inside in "data-masking" functions like `summarise()` and `mutate()`. See `vignette("colwise")` for more details.

__Usage__

```{r rcell13}
#| echo: true
#| eval: false
across(.cols = everything(), .fns = NULL, ..., .names = NULL)
```

__Arguments__

  |
 -|:--
 __.cols__| `<tidy-select>` Columns to transform. You can't select grouping columns because they are already automatically handled by the verb (i.e. `summarise()` or `mutate()`).
 __.fns__| Functions to apply to each of the selected columns. Possible values are:

- A function, e.g. `mean`.

- A purrr-style lambda, e.g. `~ mean(.x, na.rm = TRUE)`

- A list of functions/lambdas, e.g. `list(mean = mean, n_miss = ~ sum(is.na(.x))`

- `NULL`: the default value, returns the selected columns in a data frame without applying a transformation. This is useful for when you want to use a function that takes a data frame.

Within these functions you can use `cur_column()` and `cur_group()` to access the current column and grouping keys respectively.
 __.names__|  A glue specification that describes how to name the output columns. This can use `{.col}` to stand for the selected column name, and `{.fn}` to stand for the name of the function being applied. The default `NULL` is equivalent to `{.col}` for the single function case and `{.col}_{.fn}` for the case where a list is used for .fns.

### Method 1

To use the `across` in this case:

```{r rcell2}
#| echo: true
#| eval: false
test_data %>%
  mutate(across(.cols = starts_with('num_var1'),
                .fns = ~sum_c(., get(gsub('num_var1', "denom1", cur_column()))),
                .names = "{gsub('num_var1', 'var1', .col)}"))
```
```{r rcell11612sd }
#| echo: false
#| eval: true
#| message: FALSE
test_data %>%
  mutate(across(.cols = starts_with('num_var1'),
                .fns = ~sum_c(., get(gsub('num_var1', "denom1", cur_column()))),
                .names = "{gsub('num_var1', 'var1', .col)}")) %>%
knitr::kable("html")%>%
  kableExtra::kable_styling(font_size = 16)
```

Steps:

1. select referenced columns by `starts_with('num_var1')`
2. apply the function `sum_c`. here, the first argument `x` is from the `.cols`, the second argument `y` is retrived by using the function
`cur_column()`.
3. give the names to new columns

### Method 2

this code will be used for a function that can do the vectorized calculation. it's faster that the method 1.

```{r rcell21}
#| echo: true
#| eval: false
test_data %>%
  mutate(Vectorize(sum_c)(across(starts_with('num_var1'),
                                 .names = "{gsub('num_var1', 'var1', .col)}"),
                          across(starts_with('denom1')))) %>%
  do.call(data.frame, .) %>%
  rename_at(dplyr::vars(starts_with("Vectorize")), ~ gsub("Vectorize.sum_c.......", "", .))
```
```{r rcell116112sd }
#| echo: false
#| eval: true
#| message: FALSE
test_data %>%
  mutate(Vectorize(sum_c)(across(starts_with('num_var1'),
                                 .names = "{gsub('num_var1', 'var1', .col)}"),
                          across(starts_with('denom1')))) %>%
  do.call(data.frame, .) %>%
  rename_at(dplyr::vars(starts_with("Vectorize")), ~ gsub("Vectorize.sum_c.......", "", .)) %>%
knitr::kable("html")%>%
  kableExtra::kable_styling(font_size = 16)
```

Steps:

1. in the `mutate`, the function firstly is vectorized since we need a vector result returned.
2. the function `sum_c` take the first `across()` as argument `x`, the name of new columns is set in here. the second `across()`
select the needed columns as argument `y`.
3. both passed value `x`, `y`  actually are two `dataframe`. e.g. the first one is `test_data %>% select(starts_with('num_var1'))`, the second one is `test_data %>% select(starts_with('denom1'))`. to accept the `dataframe` value, the  `sum_c` need to be a vectorized function.
4. the `mutate` return a `matrix` as the new column, which can be transformed by `data.frame`(`as.data.frame` not work)