### 2023-04-24 posts/manipulate/draft_tlf_generator.qmd
Create draft_tlf_generator.qmd

---
title: 'Draft TLF generator'
date: '2023-04-19'
author: Qing Zou
categories: ['R', 'R6class']
description: 'Draft'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(checkmate)
library(R6)
library(assertive)
library(dplyr)
library(rlang)
library(Hmisc)
library(data.table)
box::use(R/utils)
```

## AdsModel

Create a `AdsModel` to manage the datasets level process

```{r rcell15}
#| echo: true
#| eval: true
#' @title AdsModel Class
#'
#'
#' @description
#' This is the  base class for ADaM dataset, to manage the CDISC attributes on dataset level
#'
#' AdsModel are build around the three following key parts:
#'
#' * Methods
#'
#'
#' For now, only implement that adding the attribute Model label by default. see `add_model_name()`
#'
#' test <- AdsModel$new()
#' test$initialize(adsl, adpsl)
#'
#' names(test$dataset)
#' a <- test$dataset
#'
#' tmp <- test$get_dataset('adsl')
#' attributes(tmp)
#'
#' adsl2 <- adsl
#' test$add_dataset(adsl2)
#' names(test$dataset)
#'
#' ```
#'

AdsModel <- R6::R6Class(
  "AdsModel",
  public = list(
    lib = NA,
    initialize = function(...){
      df_lst <-  tibble::lst(...)
      if(is.character(check_list(df_lst, min.len = 1))){
        # message("No datasets passed")
      }else{
        df_lst <- purrr::map2(df_lst, names(df_lst), ~utils$add_model_name(.x, .y))
        private$.dataset <- private$.dataset |> append(df_lst)
      }

    },
    get_dataset = function(domain){
      if(missing(domain)){
        return(private$.dataset)
      }else{
        checkmate::check_character(domain,
                                   min.chars = 1L,
                                   any.missing = FALSE)
        if(length(domain) == 1){
          return(purrr::keep_at(private$.dataset, domain)[[1]])
        }else{
          return(purrr::keep_at(private$.dataset, domain))
        }

      }
    },
    add_dataset = function(...){
        df_lst <-  tibble::lst(...)
        if(is.character(check_list(df_lst, min.len = 1))){
          message("No datasets passed")
        }else{
          df_lst <- purrr::map2(df_lst, names(df_lst), ~utils$add_model_name(.x, .y))
          private$.dataset <- private$.dataset |> append(df_lst)
        }

      },
    remove_dataset = function(...){
      # to be developed
    }
  ),
  active = list(
    dataset = function(){
        return(private$.dataset)
    }
  ),
  private = list(
    .dataset = list()
  )
)
```

```{r}
#| echo: true
#| eval: true
# test <- AdsModel$new(adsl)
adsl <- haven::read_xpt('../../data/adam/adsl.xpt')
test <- AdsModel$new()

test$initialize(adsl)

names(test$dataset)
```

```{r}
#| echo: true
#| eval: true

tmp <- test$get_dataset('adsl')
attributes(tmp)
```

```{r}
#| echo: true
#| eval: true

adtte <- haven::read_xpt('../../data/adam/adtte.xpt')
test$add_dataset(adtte)
names(test$dataset)
tmp2 <- test$get_dataset('adtte')
attributes(tmp2)
```

## VariableProc

Create a `VariableProc` to manage the variable level process

```{r rcell116}
#| echo: true
#| eval: true

VariableProc <- R6::R6Class(
  "VariableProc",
  inherit = AdsModel,
  public = list(
    # datasets level dictionary, can be retrieved by dataset
    # variables level dictionary, can be retrieved by dataset$variable
    dictionary = list(),
    initialize = function(dictionary_path = NA){

      #check dictionary_path is string
      if(!is.na(dictionary_path)){
        utils$check_dictionary_path(dictionary_path, "dictionary_path")
        dictionary <-  jsonlite::fromJSON(dictionary_path)
        self$dictionary <- dictionary
      }

    },
    get_decode = function(){
      if(is.character(check_list(self$dictionary, min.len = 1))){
        message("No dictionary avaiable, the variables will not be decoded")
      }else{
        private$.dataset <- purrr::map(private$.dataset, ~utils$set_attr(df = ., dictionary = self$dictionary))
      }

    },
    update_decode = function(dataset = NA, dictionary = ""){
      #check if the dataset exist first
      #check the format of dictionary

      # to be developed
    },
    update_detail = function(){
      # element track which dataset and variable be decoded, is it necessary integrated to S3 'summary'?

      # to be developed
    }
  ),

  #set dataset private, except variable attribute, user can't modify dataset in this module
  private = list(
    .dataset = list()
  )
)
```

```{r rcell11612 }
#| echo: true
#| eval: true
test2 <- VariableProc$new()

test2$initialize(dictionary_path = "dictionary.json")

test2$add_dataset(adsl)
adsl_tmp <- test2$get_dataset('adsl')

attributes(adsl_tmp)

attributes(adsl_tmp$SEX)$format

```

```{r}
test2$get_decode()

adsl_tmp2 <- test2$get_dataset('adsl')
attributes(adsl_tmp)
attributes(adsl_tmp2$SEX)$format

attributes(adsl_tmp2$AGEGR1)$format
```