### 2023-05-12 posts/report/data.validator.qmd
Add files via upload

---
title: 'Create a QC validation report by package data.validator'
date: '2023-05-12'
author: Qing Zou
categories: ['report', 'data.validator']
description: 'way to automated validation'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

In the `TLFgenerator`, we get two demographic example 1 and 2.

```{r lib}
#| echo: false
#| eval: true
library(dplyr)
```

```{r read1}
#| echo: true
#| eval: true
example1 <- readRDS('data/example1.rds')
```

```{r ex1}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r read2}
#| echo: true
#| eval: true
example2 <- readRDS('data/example2.rds')
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example2, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

to present the feature of data.validator, we change some value of the example output:

```{r}
#| echo: true
#| eval: true
example1_qc <- example1
example1_qc[3,1] <- "Age Group [n %]"
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example1_qc, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r}
#| echo: true
#| eval: true
example2_qc <- example2
example2_qc[23, 1] <- "^{unicode} 30"
example2_qc[18, 2] <- "15 ; 38"
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example2_qc, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r}
#| echo: true
#| eval: true
#| message: FALSE

library(assertr)
library(data.validator)

check_cols_value <- function(expected_value){
  function(x) { x %in% expected_value}
}

report <- data_validation_report()

validate(example1, name = "Verifying demographic_example1") %>%
  validate_cols(check_cols_value(example1_qc$V1), V1, description = "Check the first column V1") %>% 
  validate_cols(check_cols_value(example1_qc$V2), V2, description = "Check the first column V2") %>% 
  validate_cols(check_cols_value(example1_qc$V3), V3, description = "Check the first column V3") %>% 
  validate_cols(check_cols_value(example1_qc$V4), V4, description = "Check the first column V4") %>% 
  add_results(report)

validate(example2, name = "Verifying demographic_example2") %>%
  validate_cols(check_cols_value(example2_qc$V1), V1, description = "Check the first column V1") %>% 
  validate_cols(check_cols_value(example2_qc$V2), V2, description = "Check the first column V2") %>% 
  validate_cols(check_cols_value(example2_qc$V3), V3, description = "Check the first column V3") %>% 
  validate_cols(check_cols_value(example2_qc$V4), V4, description = "Check the first column V4") %>% 
  add_results(report)

```

```{r}
#| echo: true
#| eval: false
save_report(report, output_file = "Check_TLFs.html", output_dir = getwd())
```
<iframe src="Check_TLFs.html" height="100%" width="100%"></iframe>